### 2021-08-31 paper.Rmd
huber plots

When using a tour, a sequence of $d$-dimensional linear projections are obtained from a $p$-dimensional space, where $d \ll p$. Let $X_{n \times p}$ be the data matrix, consisting of $n$ observations and $p$ variables, whose projections are of interest. A projection basis, $A_{p \times d}$, is a matrix that characterizes the direction from which the data are projected.  It needs to satisfy an orthonormality condition, which requires that all the columns in $A$ has unit length and are perpendicular to each other. With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$. The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure. Two variables and the species labels from the palmer penguins data [@penguins] are used for illustration. In the Huber plots (left) a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of illustrates the projection shows two examples of low-dimensional projections of the palmer penguins data [@penguins] in a Huber plot [@huber1990] and a histogram. (The figures produced with the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.)

>>>>

When using a tour, a sequence of $d$-dimensional linear projections are obtained from a $p$-dimensional space, where $d \ll p$. Let $X_{n \times p}$ be the data matrix, consisting of $n$ observations and $p$ variables, whose projections are of interest. A projection basis, $A_{p \times d}$, is a matrix that characterizes the direction from which the data are projected.  It needs to satisfy an orthonormality condition, which requires that all the columns in $A$ has unit length and are perpendicular to each other. With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$. The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure. (The figures produced with the R package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration. In the Huber plots (left) [@huber1990] a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of a PP index, computed for all 1D projections. (The dashed circle is a guide to help perceive high and low PP index values.) The line segments indicate the 1D projection which is plotted in the histogram to the right. The histogram exhibits in the top row clustering of the three species, and correspond to the 1D projection with a high PP index value. The histogram in the bottom row corresponds to a 1D projection with low PP index value, and can be seen to have no interesting pattern.
